---
title: "Reproducable Research"
output: html_document
keep_md: true
---
# Reproducable Research: Assignment 1


##Loading the data
##### Load the data and dependencies

```{r}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)

fulldata <- read.csv("activity.csv")
```

------

##What is the mean total number of steps taken per day?
  
##### Calculate the steps per day

```{r}
library(ggplot2)
library(dplyr)

dailysteps <- tapply(fulldata$steps, fulldata$date, sum, na.rm = TRUE)
```

##### Create a histogram
```{r}
library(ggplot2)

qplot(dailysteps, xlab = 'Total number of steps per day', ylab = 'Frequency using binwidth 500')

```

##### Mean and median calculations

```{r}
mean_steps <- mean(dailysteps, na.rm = TRUE)

## [1] 9354.23
```

```{r}
median_steps <- median(dailysteps, na.rm = TRUE)

##[1] 10395
```


## What is the average daily pattern?


```{r}
avgdailypattern <- aggregate(x=list(steps = fulldata$steps), by =  list(interval = fulldata$interval), FUN = mean, na.rm = TRUE)

ggplot(data = avgdailypattern, aes(x = interval, y = steps)) +
  geom_line() +
  xlab("5-minute interval") +
  ylab("average number of steps taken")

```

##### On average across all of the days in the dataset, the 5 minute interval contains the maximum number of steps?

```{r}
maxsteps <- filter(avgdailypattern, steps == max(steps))
```

```{r}
##     interval    steps
##      835    206.1698
```

##Imputing missing values

```{r}
missing_values <- sum(is.na(fulldata$steps))

##[1] 2304
```

```{r}
fulldataImputed <- fulldata


mean_steps <- mean(fulldata$steps, na.rm = TRUE)


fulldataImputed$steps[is.na(fulldataImputed$steps)] <- mean_steps

```

```{r}
steps_imputed_by_day <- tapply(fulldataImputed$steps, fulldataImputed$date, sum)

qplot(steps_imputed_by_day, xlab = 'Total Num. Imputed Steps', ylab = 'Frequency Using Binwith 500', binwidth = 500)



```

```{r}
meanImputed <- mean(steps_imputed_by_day)

medianImputed <- median(steps_imputed_by_day)

##[1] 10766.19
##[1] 10766.19
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lubridate)


fulldataImputed <- fulldata


fulldataImputed$date <- as.POSIXlt(fulldataImputed$date)

fulldataImputed$dateType <- ifelse(weekdays(fulldataImputed$date) %in% c('Saturday', 'Sunday'), 'weekend', 'weekday')


```

```{r}
averagedfulldataImputed <- aggregate(steps ~ interval + dateType, data = fulldataImputed, mean)

ggplot(averagedfulldataImputed, aes(interval, steps)) +
  geom_line() + 
  facet_grid(dateType ~ .) +
  xlab("5-minute interval") + 
  ylab("average number of steps")

```


